---
title: "Handling non-standard weekly data in R"
subtitle: "An overview of 'aweek', the lightweight R package for handling weekly data"
author: "Zhian N. Kamvar"
output: 
  md_document:
    variant: markdown_github
    preserve_yaml: true
---

This week (`r aweek::date2week('2019-06-11', week_start = 1)`), the [*aweek* package](../aweek) 
version 1.0.0 has been [released on 
CRAN](https://cran.r-project.org/package=aweek).

The *aweek* package is a response to the lack of utilities to easily define
weeks that do not start on Monday, which is necessary for outbreak response in
[countries whose calendars may start on Sunday or
Saturday](https://en.wikipedia.org/wiki/File:First_Day_of_Week_World_Map.svg).
This package provides utilities to convert dates to weeks and back again with
any week start definition.

You can install *aweek* with:

```{r, eval = FALSE}
install.packages("aweek")
```

Using aweek is straightforward. When you load the package, you can set the 
day on which all weeks should start with `set_week_start()`.

```{r}
library("aweek")
# At the beginning of your analysis, set the week start to the weeks you want
# to use for reporting
ow <- set_week_start("Sunday")
d  <- as.Date("2019-06-11")
(w <- as.aweek(d)) # convert dates to weeks
as.Date(w)         # convert weeks to dates
```

All aweek objects carry a `week_start` attribute, which is the ISO week day that
defines the start of the week for that object.

One of the common things you might want to do is to define a range of dates to
use to filter linelist data. You can use `get_date()` to get a date from week
numbers. Note that day in this context refers to the day of the week relative to
the week start, so a "7" is "Saturday".

```{r get_date}
library("dplyr")
library("tidyr")
linelist   <- as_tibble(outbreaks::ebola_sim_clean$linelist)
linelist
(obs_start <- get_date(week = 10, year = 2014, day = 1))
(obs_end   <- get_date(week = 40, year = 2014, day = 7))

ll_obs <- linelist %>%
  filter(date_of_onset >= obs_start) %>%
  filter(date_of_onset <= obs_end) %>%
  filter(hospital == "other")
ll_obs
```

You can use aweek to aggregate weekly data. The default aweek object is in the
ISO 8601 format (YYYY-Www-d) and `factor_aweek()` can remove the `-d` portion,
giving weekly aggregates, which will include any missing weeks in the levels:

```{r aggregation}
ll_obs %>%
  mutate(week = as.aweek(date_of_onset)) %>%
  mutate(aggregate_week = factor_aweek(week)) %>%
  count(aggregate_week, gender, .drop = FALSE) %>%
  spread(gender, n)
```


