<!DOCTYPE html>
<html lang="en">
  <!-- Beautiful Jekyll | MIT license | Copyright Dean Attali 2016 -->
  <head>
  <meta charset="utf-8" />
  <meta http-equiv="X-UA-Compatible" content="IE=edge">
  <meta name="viewport" content="width=device-width, initial-scale=1.0, maximum-scale=1.0">

  <title>Handling non-standard weekly data in R</title>

  <meta name="author" content="RECON - R Epidemics Consortium" />
  
  
  <meta name="description" content="An overview of 'aweek', the lightweight R package for handling weekly data">
  

  <link rel="alternate" type="application/rss+xml" title="R Epidemics Consortium - We are a group of international experts in infectious disease modelling, Public Health, and software development who collaborate to create the next generation of analysis tools for disease outbreak response using the R software" href="/feed.xml" />

  
    
      <link rel="stylesheet" href="//maxcdn.bootstrapcdn.com/font-awesome/4.5.0/css/font-awesome.min.css" />
    
      
  
  
    
      <link rel="stylesheet" href="/css/bootstrap.min.css" />
    
      <link rel="stylesheet" href="/css/bootstrap-social.css" />
    
      <link rel="stylesheet" href="/css/main.css" />
    
      <link rel="stylesheet" href="/css/recon.css" />
    
    
  
  
    
      <link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lora:400,700,400italic,700italic" />
    
      <link rel="stylesheet" href="//fonts.googleapis.com/css?family=Open+Sans:300italic,400italic,600italic,700italic,800italic,400,300,600,700,800" />
    
  

    
  
  

  

  <!-- Facebook OpenGraph tags -->
  <meta property="og:title" content="Handling non-standard weekly data in R" />
  <meta property="og:type" content="website" />
  
  <meta property="og:url" content="http://reconhub.github.io/posts-Rmd-files/2019-06-12-aweek-1.0.0.Rmd" />
  
  
  <meta property="og:image" content="" />
  
  
  <!-- Twitter summary cards -->
  <meta name="twitter:card" content="summary" />
  <meta name="twitter:site" content="@yourname" />
  <meta name="twitter:creator" content="@yourname" />
  <meta name="twitter:title" content="Handling non-standard weekly data in R" />
  <meta name="twitter:description" content="This week (`r aweek::date2week(&apos;2019-06-11&apos;, week_start = 1)`), the [*aweek* package](../aweek) version 1.0.0 has been [released on CRAN](https://cran.r-project.org/package=aweek). The *aweek* package is a response to the lack of utilities to easily define weeks that do not start on Monday, which is necessary for outbreak response in [countries whose calendars may start..." />
  
  
</head>


  <body>
  
    <nav class="navbar navbar-default navbar-fixed-top navbar-custom">
  <div class="container-fluid">
    <div class="navbar-header">
      <button type="button" class="navbar-toggle" data-toggle="collapse" data-target="#main-navbar">
        <span class="sr-only">Toggle navigation</span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
        <span class="icon-bar"></span>
      </button>
      <a class="navbar-brand" href="http://reconhub.github.io">R Epidemics Consortium</a>
    </div>

    <div class="collapse navbar-collapse" id="main-navbar">
      <ul class="nav navbar-nav navbar-right">
      
        
          <li class="navlinks-container">
            <a class="navlinks-parent" href="javascript:void(0)">About RECON</a>
            <div class="navlinks-children">
              
                
                  
            





<a href="/">What we do</a>

                
              
                
                  
            





<a href="/join">Join us</a>

                
              
                
                  
            





<a href="/contact">Contact</a>

                
              
                
                  
            





<a href="/CODE_OF_CONDUCT">Code of Conduct</a>

                
              
            </div>
          </li>
        
        
        
          <li>
            
            





<a href="/blog">News</a>

          </li>
        
        
        
          <li>
            
            





<a href="/people">People</a>

          </li>
        
        
        
          <li>
            
            





<a href="/projects">Projects</a>

          </li>
        
        
        
          <li>
            
            





<a href="/resources">Resources</a>

          </li>
        
        
        
          <li>
            
            





<a href="/events">Events</a>

          </li>
        
        
        
          <li>
            
            





<a href="/funding">Funding</a>

          </li>
        
        
        
          <li>
            
            





<a href="/forum">Forum</a>

          </li>
        
        
      </ul>
    </div>

	

  </div>
</nav>


    <!-- TODO this file has become a mess, refactor it -->





<header class="header-section ">

<div class="intro-header no-img">
  <div class="container">
    <div class="row">
      <div class="col-lg-8 col-lg-offset-2 col-md-10 col-md-offset-1">
        <div class="page-heading">
          <h1>Handling non-standard weekly data in R</h1>
		  
		    
            <hr class="small">
            <span class="page-subheading">An overview of 'aweek', the lightweight R package for handling weekly data</span>
			
		  
		  
		  
        </div>
      </div>
    </div>
  </div>
</div>
</header>





<div class="container" role="main">
  <div class="row">
    <div class="col-lg-8 col-lg-offset-2 col-md-10 col-md-offset-1">
      This week (`r aweek::date2week('2019-06-11', week_start = 1)`), the [*aweek* package](../aweek) 
version 1.0.0 has been [released on 
CRAN](https://cran.r-project.org/package=aweek).

The *aweek* package is a response to the lack of utilities to easily define
weeks that do not start on Monday, which is necessary for outbreak response in
[countries whose calendars may start on Sunday or
Saturday](https://en.wikipedia.org/wiki/File:First_Day_of_Week_World_Map.svg).
This package provides utilities to convert dates to weeks and back again with
any week start definition.

You can install *aweek* with:

```{r, eval = FALSE}
install.packages("aweek")
```

Using aweek is straightforward. When you load the package, you can set the 
day on which all weeks should start with `set_week_start()`.

```{r}
library("aweek")
# At the beginning of your analysis, set the week start to the weeks you want
# to use for reporting
ow <- set_week_start("Sunday")
d  <- as.Date("2019-06-11")
(w <- as.aweek(d)) # convert dates to weeks
as.Date(w)         # convert weeks to dates
```

All aweek objects carry a `week_start` attribute, which is the ISO week day that
defines the start of the week for that object.

One of the common things you might want to do is to define a range of dates to
use to filter linelist data. You can use `get_date()` to get a date from week
numbers. Note that day in this context refers to the day of the week relative to
the week start, so a "7" is "Saturday".

```{r get_date}
library("dplyr")
library("tidyr")
linelist   <- as_tibble(outbreaks::ebola_sim_clean$linelist)
linelist
(obs_start <- get_date(week = 10, year = 2014, day = 1))
(obs_end   <- get_date(week = 40, year = 2014, day = 7))

ll_obs <- linelist %>%
  filter(date_of_onset >= obs_start) %>%
  filter(date_of_onset <= obs_end) %>%
  filter(hospital == "other")
ll_obs
```

You can use aweek to aggregate weekly data. The default aweek object is in the
ISO 8601 format (YYYY-Www-d) and `factor_aweek()` can remove the `-d` portion,
giving weekly aggregates, which will include any missing weeks in the levels:

```{r aggregation}
ll_obs %>%
  mutate(week = as.aweek(date_of_onset)) %>%
  mutate(aggregate_week = factor_aweek(week)) %>%
  count(aggregate_week, gender, .drop = FALSE) %>%
  spread(gender, n)
```



	    
    </div>
  </div>
</div>


    <footer>
  <div class="container beautiful-jekyll-footer">
    <div class="row">
      <div class="col-lg-8 col-lg-offset-2 col-md-10 col-md-offset-1">
        <ul class="list-inline text-center footer-links">
		    
          <li>
            <a href="mailto:thibautjombart@gmail.com" title="Email me">
              <span class="fa-stack fa-lg">
                <i class="fa fa-circle fa-stack-2x"></i>
                <i class="fa fa-envelope fa-stack-1x fa-inverse"></i>
              </span>
            </a>
          </li>
        
        
          <li>
            <a href="https://github.com/reconhub/reconhub.github.io" title="GitHub">
              <span class="fa-stack fa-lg">
                <i class="fa fa-circle fa-stack-2x"></i>
                <i class="fa fa-github fa-stack-1x fa-inverse"></i>
              </span>
            </a>
          </li>
        
          

		  
      
		  
        </ul>
        <p class="copyright text-muted">
		  RECON - R Epidemics Consortium
		  &nbsp;&bull;&nbsp;
		  2022
	    </p>
		<p class="theme-by text-muted">
		  Theme by
		  <a href="http://deanattali.com/">Dean Attali</a>
		</p>
      </div>
    </div>
  </div>
</footer>

  
    






  
    <!-- doing something a bit funky here because I want to be careful not to include JQuery twice! -->
    
      <script>
      	if (typeof jQuery == 'undefined') {
      	  document.write('<script src="/js/jquery-1.11.2.min.js"></scr' + 'ipt>');
      	}
      </script>
    
  
    <!-- doing something a bit funky here because I want to be careful not to include JQuery twice! -->
    
	<script src="/js/bootstrap.min.js"></script>
    
  
    <!-- doing something a bit funky here because I want to be careful not to include JQuery twice! -->
    
	<script src="/js/main.js"></script>
    
  



	<!-- Google Analytics -->
	<script>
		(function(i,s,o,g,r,a,m){i['GoogleAnalyticsObject']=r;i[r]=i[r]||function(){
		(i[r].q=i[r].q||[]).push(arguments)},i[r].l=1*new Date();a=s.createElement(o),
		m=s.getElementsByTagName(o)[0];a.async=1;a.src=g;m.parentNode.insertBefore(a,m)
		})(window,document,'script','//www.google-analytics.com/analytics.js','ga');
		ga('create', 'UA-83713279-1', 'auto');
		ga('send', 'pageview');
	</script>
	<!-- End Google Analytics -->


  
  </body>
</html>
